\documentclass[12pt,a4paper,oneside,titlepage,abstract]{scrartcl} 
\usepackage[T1]{fontenc}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[utf8]{inputenc}
\usepackage[italian,english]{babel} %vedi arte pag.20 in caso si decida di inserire grossa citazione inglese
\usepackage[italian]{varioref}		% per riferimenti più completi con \vref
\usepackage{microtype}				% migliora il riempimento delle righe
% tramite espansione del font e protrusione
% dei caratteri (ars pag.49)

\usepackage[autostyle,italian=guillemets]{csquotes}

\usepackage[backend=biber,bibstyle=ieee,citestyle=numeric-comp]{biblatex}
\addbibresource{bibliography.bib}

\usepackage{quoting}				% per citazione in "quote"
\quotingsetup{font=small} % ,font=itshape

%Tabelle
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{caption}
\captionsetup{position=above, font=small, format=hang, labelfont={sf,bf}} % versione adattata a komascript
%\captionsetup{tableposition=top,figureposition=bottom,font=small,format=hang,labelfont={sf,bf}} versione generica/precedente
\setkomafont{captionlabel}{\sffamily\bfseries}
\setkomafont{caption}{\small}
\captionsetup[table]{position=top}
\captionsetup[figure]{position=bottom}

%\setkomafont{caption}{\itshape\sffamily}
%\setkomafont{captionlabel}{\upshape\bfseries\sffamily}

\newcolumntype{Y}{>{\centering\arraybackslash}X}
\usepackage{multirow}
\usepackage{subfig}

%\usepackage{fancyhdr}

\usepackage{amsmath}				% AMS = American Mathematical Society
\usepackage{amssymb}				% per simboli matematici
\usepackage{mathtools}				% altri tool, basato su amsmath, vedi "definizioni.sty"
\usepackage{centernot}				% per negare simboli matematici sovrapponendo la barra al centro
\usepackage{braket}					% utile per gli insiemi, definisce \Set
\usepackage{amsthm}					% per enunciati e dimostrazioni

\usepackage[output-decimal-marker={.}]{siunitx} % Per le unità di misura 
% internazionali e la virgola come separatore
% decimale, contiene \num per scrivere
% correttamente i numeri (virgola e spazi)

\newcommand{\omissis}{[\textellipsis\unkern]} % definizione del comando per inserire [...]
\usepackage{eurosym}

\usepackage{definizioni}

\usepackage{hyperref}%[colorlinks,linkcolor=blue]
\hypersetup{% Esempio di opzioni per hyperref
	pdfpagemode={UseOutlines},
	bookmarksopen,
	pdfstartview={FitH},
	colorlinks,
	linkcolor={blue},
	citecolor={blue},
	urlcolor={blue}
}
% per collegamenti ipertestuali, va caricato
% per ultimo. Permette collegamenti al web
% con \href e carica automaticamente anche
% il pacchetto url (che include \url)

\begin{document}
	\titlehead{RLSB}
	\subject{Software Engineering for Artificial Intelligence}
	\title{Reinforcement Learning Sustainability Benchmark}
	\subtitle{Report v1.0}
	\author{Luca Strefezza\thanks{l.strefezza1@studenti.unisa.it}}
	%\date{Data da inserire}
	\maketitle
	
\begin{abstract}
	This study presents a comprehensive benchmark of the energy consumption and performance trade-offs of deep reinforcement learning (DRL) algorithms, providing novel insights into their sustainability implications. Although advancements in DRL have traditionally prioritized performance, the associated computational costs and environmental impacts remain underexplored.
	
	We systematically evaluate several value-based and policy gradient methods—including DQN, Double DQN, Prioritized Experience Replay, Dueling DQN, C51, PPO, REINFORCE, and SAC—using a \num{100000}-step Atari benchmark. Each algorithm is rigorously assessed across \num{8} Atari games with \num{4} random-seeded runs per game to ensure statistically robust comparisons. To capture detailed performance and energy metrics, we employ tools such as Weights and Biases, TensorBoard, and CodeCarbon within a development environment based on CleanRL, which guarantees consistency and reproducibility.
	
	Key findings indicate that Proximal Policy Optimization (PPO) achieves the highest energy efficiency (\num{0.0029} kgCO\textsubscript{2}eq), whereas Soft Actor-Critic (SAC) records the highest energy consumption (\num{0.015} kgCO\textsubscript{2}eq) despite strong performance in select environments. DQN-based methods strike a balance, delivering moderate emissions (\num{0.006}--\num{0.008} kgCO\textsubscript{2}eq) alongside varied game-specific advantages. Additionally, our results suggest that short-horizon constraints can limit the effectiveness of more complex architectures, such as those inspired by Rainbow.
	
	By providing empirical evidence on the energy-performance trade-offs in DRL, this study contributes to the growing discourse on sustainable AI and underscores the need for balancing computational efficiency with algorithmic advancements.
\end{abstract}

	
	{
		\hypersetup{linkcolor=black}
		\tableofcontents
		\listoftables
		\listoffigures
	}
	
	\input{sections/context}
	\input{sections/goals}
	\input{sections/methodologies}
	\input{sections/preliminary_results}
	\input{sections/implications_results}
	\input{sections/conclusions}
		
	\printbibliography[heading=bibintoc]
	
\end{document}
